---
title: 'Assignment 5: Hierarchical Clustering'
author: "Jared Bartee"
date: "`r Sys.Date()`"
output:
  word_document: default
  html_document: default
---
# Step 1: Data Preprocessing
# Load the data
```{r}
cereals_data <- read.csv("Cereals.csv")
head(cereals_data)
```
# Drop mfr & type columns
```{r}
cereals_data <- cereals_data[-c(2,3)]
cereals_data
```
# Convert the names of breakfast cereals to row names
```{r}
rownames(cereals_data) <- cereals_data$name
```
#Drop the name column
```{r}
cereals_data <- cereals_data[-c(1)]
cereals_data
```
# Find the amount of missing values and then omit them
```{r}
sum(is.na(cereals_data))
```

```{r}
cereals_data <- na.omit(cereals_data)
sum(is.na(cereals_data))
```

# Step 2: Hierarchical Clustering
# Normalize the data
```{r}
normalized_cereals <- scale(cereals_data[, -1])
```

# Use the Euclidean distance measure
```{r}
cereals_euclidean <- dist(normalized_cereals, method = "euclidean")
plot(cereals_euclidean)
```
```{r}
CC1 <- hclust(cereals_euclidean)
plot(CC1, cex = 0.5, hang = -1)
```
# Use Agnes
```{r}
library(cluster)
library(stats)
Cereal_Single <- agnes(normalized_cereals, method = "single")
Cereal_Complete <- agnes(normalized_cereals, method = "complete")
Cereal_average <- agnes(normalized_cereals, method = "average")
Cereal_Ward <- agnes(normalized_cereals, method = "ward")
```

```{r}
print(Cereal_Single$ac)
```

```{r}
print(Cereal_Complete$ac)
```

```{r}
print(Cereal_average$ac)
```
# Ward appears to be the best method
```{r}
print(Cereal_Ward$ac)
```

# Apply hierarchical clustering with Ward's method & Visualize for Ward's method
```{r, fig.height=8, fig.width=14}
pltree(Cereal_Ward,main = "Dendrogram of agnes: Ward")
```
```{r, fig.height=8, fig.width=14}
plot(Cereal_Ward)
rect.hclust(Cereal_Ward, k = 5, border = 1:5) # I would choose 5 clusters.
abline(h = 5, col = 'red')
```

# Step 3: Cluster Stability and Healthy Cereals
# Create cluster partitions A and B
```{r}
set.seed(123)
partition_A <- sample(1:2, nrow(normalized_cereals), replace = TRUE)
partition_B <- 3 - partition_A
```

# Fit cluster on partition A
```{r}
cluster_A <- cutree(Cereal_Ward, k = 5)  # Assuming 4 clusters, adjust as needed
```

# Calculate centroids of clusters in partition A
```{r}
library(dplyr)
```

# Combine partition_A, cluster_A, and normalized_cereals for easier processing
```{r}
combined_data <- cbind(partition = partition_A, cluster = cluster_A)
combined_data_df <- as.data.frame(combined_data)
```

# Calculate centroids of clusters in partition A
```{r}
cluster_A_centroids <- combined_data_df %>%
  group_by(cluster) %>%
  summarise_all(mean) %>%
  select(-partition)
```

# Calculate distances between partition B records and centroids from partition A
```{r}
distances_to_centroids <- apply(normalized_cereals[partition_B, ], 1, function(record) {
  apply(cluster_A_centroids, 1, function(centroid) {
    sqrt(sum((record - centroid)^2))
  })
})
```

# Assign clusters to partition B based on minimum distances
```{r}
cluster_B <- apply(distances_to_centroids, 2, which.min)
```


# Assess cluster consistency
```{r}
consistency <- sum(cluster_A == cluster_B) / length(cluster_B)
```

# Identify healthy cereals cluster
# Let's assume 'healthy' cereals have low sugar and high fiber
```{r}
healthy_cereals_cluster <- cluster_A[which(cereals_data$sugars < 5 & cereals_data$fiber > 5)]
```

# Print the results
```{r}
cat("Cluster Consistency:", consistency, "\n")
cat("Healthy Cereals Cluster:", healthy_cereals_cluster, "\n")
```










