---
title: "Assignment 3"
author: "Jared Bartee"
date: "`r Sys.Date()`"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
library("dplyr")
library("tidyr")
library("ggplot2")
library("ROCR")
library("rpart")
library("rpart.plot")
library("caret")
library("randomForest")
library("tidyverse")
library("tm")
library("SnowballC")
library("softImpute")
library("glmnet")
library("Hmisc")
library("dummy")
library('tinytex')
library('GGally')
library('gplots')
library('FNN')
library("dplyr")
library("tidyr")
library("caTools")
library("ggpubr")
library("reshape2")
library("e1071")
```


```{r}
bank = read.csv("UniversalBank.csv")
summary(bank)
```
```{r}
bank$Personal.Loan = as.factor(bank$Personal.Loan)
bank$Online = as.factor(bank$Online)
bank$CreditCard = as.factor(bank$CreditCard)
set.seed(1)
train.index <- sample(row.names(bank), 0.6*dim(bank)[1])
test.index <- setdiff(row.names(bank), train.index)
train.df <- bank[train.index, ]
test.df <- bank[test.index, ]
train <- bank[train.index, ]
test = bank[train.index,]
```


#A Create a pivot table for the training data with Online as a column variable, CC as a row variable, and Loan as a secondary row variable. The values inside the table should convey the count. In R use functions melt() and cast(), or function table(). In Python, use panda dataframe methods melt() and pivot().
```{r}
melted.bank = melt(train,id=c("CreditCard","Personal.Loan"),variable= "Online")
recast.bank=dcast(melted.bank,CreditCard+Personal.Loan~Online)
recast.bank[,c(1:2,14)]
```

#B  Consider the task of classifying a customer who owns a bank credit card and is actively using online banking services. Looking at the pivot table, what is the probability that this customer will accept the loan offer? [This is the probability of loan acceptance (Loan = 1) conditional on having a bank credit card (CC = 1) and being an active user of online banking services (Online = 1)].

Answer: Probability of Loan acceptance give having a bank credit card and user of online services is 77/3000 = 2.6%

#C  Create two separate pivot tables for the training data. One will have Loan (rows) as a function of Online (columns) and the other will have Loan (rows) as a function of CC.
```{r}
melted.bankc1 = melt(train,id=c("Personal.Loan"),variable = "Online")
```

```{r}
melted.bankc2 = melt(train,id=c("CreditCard"),variable = "Online")
```

```{r}
recast.bankc1=dcast(melted.bankc1,Personal.Loan~Online)
recast.bankc2=dcast(melted.bankc2,CreditCard~Online)
Loanline=recast.bankc1[,c(1,13)]
LoanCC=recast.bankc2[,c(1,14)]

Loanline
```
```{r}
LoanCC
```
#D Compute the following quantities [P(A | B) means “the probability ofA given B”]: i. P(CC = 1 | Loan = 1) (the proportion of credit card holders among the loan acceptors);ii. P(Online = 1 | Loan = 1);iii. P(Loan = 1) (the proportion of loan acceptors);iv. P(CC = 1 | Loan = 0);v. P(Online = 1 | Loan = 0);vi. P(Loan = 0)
```{r}
table(train[,c(14,10)])
```
```{r}
table(train[,c(13,10)])
```
```{r}
table(train[,c(10)])
```
Answers:
i. 77/(77+198)= 28%
ii. 166/166+109)= 60.3%
iii. 275/(275+2725)= 9.2%
iv. 801/(801+1924)= 29.4%
v. 1588/(1588+1137)= 58.3%
vi. 2725/(2725+275)= 90.8%

#E Use the quantities computed above to compute the naive Bayes probability P(Loan = 1 | CC = 1, Online = 1).
```{r}
((77/(77+198))*(166/(166+109))*(275/(275+2725)))/(((77/(77+198))*(166/(166+109))*(275/(275+2725)))+((801/(801+1924))*(1588/(1588+1137))*2725/(2725+275)))
```
#F Compare this value with the one obtained from the pivot table in (B). Which is a more accurate estimate?

Answer: 9.05% is very similar to the 9.7% the difference between the exact method and the naive-baize method is the exact method would need the exact same independent variable classifications to predict, whereas the naive Bayes method does not.

#G Which of the entries in this table are needed for computing P(Loan = 1 | CC = 1, Online = 1)? Run naive Bayes on the data. Examine the model output on training data, and find the entry that corresponds to P(Loan = 1 | CC = 1, Online = 1). Compare this to the number you obtained in (E).
```{r}
naive.train = train.df[,c(10,13:14)]
naive.test = test.df[,c(10,13:14)]
naivebayes = naiveBayes(Personal.Loan~.,data=naive.train)
naivebayes
```
Answer: The naive Bayes is the exact same output we received in the previous methods. (.280)(.603)(.09)/(.280.603.09+.29.58.908) = .09 which is the same response provided as above.



